An algorithm for a spatial mapping using a hexagon tilegram, with application to Australia disease maps

### 2019-01-10 algorithm_radix.Rmd
more steps in filtering the grid

The algorithm operates on a set of polygons.
There are parameters considered that may be provided, or will be automatically derived.
All necessary functions are exported, with a main function `create_hexmap` used to step through automatically.

@@

```{r}
hexmap_allocation <- allocate(centroids = centroids,
        hex_grid = grid,
        hex_size = hex_size,
        filter_dist = filter_dist,
        width = width,
        focal_points = focal_points,
        verbose = TRUE,
        id = sf_id)
```

Allocation of all centroids can now take place using the set of polygon centroids and the hexagon map grid. 
For each polygon centroid, only the hexagon grid points that have not yet been used can be considered.
The filter distance parameter is used to subset possible grid points to only those surrounding the polygon centroid within the filter distance.
Using the angle between each polygon centroid, and it's the closest focal point, the subset of points is filtered again.
Of these possible points, only those within a specific amount of degrees plus and minus a specific angle range are kept. This angle begins at 30 degrees by default, and may increase if necessary.

>>>>

The algorithm operates on a set of polygons.
There are parameters needed that may be provided, or will be automatically derived.
All necessary functions are exported, with a main function `create_hexmap` used to step through automatically.

@@

```{r, eval = FALSE}
hexmap_allocation <- allocate(centroids = centroids,
        hex_grid = grid,
        hex_size = hex_size,
        filter_dist = filter_dist,
        width = width,
        focal_points = focal_points,
        verbose = TRUE,
        id = sf_id)
```

Allocation of all centroids can now take place using the set of polygon centroids and the hexagon map grid. 
For each polygon centroid, only the hexagon grid points that have not yet been used can be considered.

```{r}
s_centroids <- centroids %>% arrange(focal_distance)

s_centroids <- split(s_centroids, s_centroids[["focal_distance"]])

# Set up allocation data frame
centroid_allocation <- NULL
```

The filter distance parameter is used to subset possible grid points to only those surrounding the polygon centroid within the filter distance.

```{r}
# keep value to reset expanded distances
expand_dist <- filter_dist
```

The following example considers one of the Local Government Areas. 
These steps are repeated for each polygon.

### Step 1: Filter the grid for unassigned hexagon points

```{r}
# filter for only the available hex grid points
hex_grid <- grid %>% filter(!assigned)
```

### Step 2: Filter the grid points for those closest to the centroid
```{r}

# filter grid for avaiable points
centroid1 <- centroids %>% head(1)

flong <- centroid1$longitude
flat <- centroid1$latitude

hex_grid <- hex_grid %>% ungroup() %>%
        filter(flat - filter_dist < hex_lat & hex_lat < flat + filter_dist) %>%
        filter(flong - filter_dist < hex_long & hex_long < flong + filter_dist)

```

Create a box filter
```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "green") + geom_point(aes(x=longitude, y = latitude), colour = "red", data= centroids[1,]) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "grey", data = hex_grid, size = 0.75) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "orange", data = hex_grid, size = 0.75)
```

```{r}
    grid <- grid %>%
        rowwise %>%
        mutate(
        hex_lat_c = hex_lat - flat,
        hex_long_c = hex_long - flong) %>%
        mutate(hyp = ((hex_lat_c^2) + (hex_long_c^2))^(1/2))

```

Create circle filter

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "green") + geom_point(aes(x=longitude, y = latitude), colour = "red", data= centroids[1,]) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "grey", data = hex_grid, size = 0.75) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "orange", data = hex_grid, size = 0.75)
```

```{r}

    # Filter for angle within circle
    if ("focal_distance" %in% colnames(centroid1)) {

        f_angle <- centroid1 %>%
            mutate(atan = atan2(latitude-latitude1,longitude-longitude1),
                angle = (atan*180/pi),
                pangle = ifelse(angle<0, angle +360, angle)) %>% pull()

        grid <- grid %>%
            # create circle of radius: filter_dist
            filter(hyp < filter_dist) %>%
            mutate(
                # geosphere takes a long time
                angle = f_angle,
                angle_plus = (angle + angle_width)%%360,
                angle_minus = (angle - angle_width)%%360,
                atan = atan2(hex_lat_c, hex_long_c),
                hex_angle = (atan*180/pi),
                hex_angle = ifelse(hex_angle<0, hex_angle +360, hex_angle))

        if (grid$angle_minus[1] < grid$angle_plus[1]) {
            grid <- grid %>%
                # create slice of 60 degrees from centroid
                filter(angle_minus < hex_angle & hex_angle < angle_plus)
        }
        else {
            grid <- grid %>%
                # create slice of 60 degrees from centroid
                filter(hex_angle < angle_plus | angle_minus > hex_angle)
        }

    }

```

```{r}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "green") + geom_point(aes(x=longitude, y = latitude), colour = "red", data= centroids[1,]) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "grey", data = hex_grid, size = 0.75) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "orange", data = hex_grid, size = 0.75)
```

```{r}

# Choose first available point
        cent <- centroid1 %>% dplyr::rename(focal_point = points, focal_dist = focal_distance, focal_angle = angle)

        # Filter should give one hex point
        hex <- hex_grid %>% 
          ungroup %>% 
          filter(hyp == min(hyp)) %>%
          select(hex_long, hex_lat, hex_id = id)

        #update grid to show this centroid as assigned
        hex_grid[which(hex_grid$id == hex$hex_id),]$assigned <- TRUE

        centroid_allocation <- bind_rows(centroid_allocation, dplyr::bind_cols(cent, hex)) %>% as.tibble()

```

Using the angle between each polygon centroid, and it's the closest focal point, the subset of points is filtered again.
Of these possible points, only those within a specific amount of degrees plus and minus a specific angle range are kept. This angle begins at 30 degrees by default, and may increase if necessary.
